\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-09-09 NAR_DRAFT.tex
add note comment
numbered citations
changes to discussion

\usepackage{NAR-natbib}
\bibliographystyle{unsrtnat}
\usepackage{color}
\usepackage{xfrac}
\newcommand{\rngcomment}[1]{{\color{red}RNG: #1}}
\newcommand{\bkmcomment}[1]{{\color{blue}BKM: #1}}
\newcommand{\batcave}{BATCAVE }
\newcommand{\beginsupplement}{%
        \clearpage
        \onecolumn
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
% Enter dates of publication
\copyrightyear{2008}
\pubdate{31 July 2009}
\pubyear{2009}
\jvolume{37}
\jissue{12}

@@

\bkmcomment{can I work a citation in here? using data sets from the same data generating distribution as a prior? Struggling to find one, but will keep looking.}
\batcave approaches the set of variant calls returned from MuTect as two distinct sets of data.
The first is the set of mutations that pass all heuristics filters, and  have a very large posterior probability of being present given the observed read data.
This data allows us to reliably estimate the distribution of substitution types, i.e. the mutation profile.
Variants with low probability given the read data are noisy measurements of samples from the same data generating distribution.
We use the estimated distribution of substitution types to adjust the prior probability of mutation for these lower probability variants.

@@

\section{CONCLUSION}
\bkmcomment{I looked at a bunch of NAR methods papers, and none include a conclusions section, so I moved it to the last paragraph of discussion.}

\section{Software Availability}
\bkmcomment{working on this today}
github for source code
zenodo for figures, simulations
zenodo for package at time of bioarxiv

@@

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text.

>>>>

\usepackage{NAR-natbib}
\bibliographystyle{NAR-natbib}
\usepackage{color}
\usepackage{xfrac}
\newcommand{\rngcomment}[1]{{\color{red}RNG: #1}}
\newcommand{\bkmcomment}[1]{{\color{blue}BKM: #1}}
\newcommand{\batcave}{BATCAVE }
\newcommand{\beginsupplement}{%
        \clearpage
        \onecolumn
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
% Enter dates of publication
\copyrightyear{2008}
\pubdate{31 July 2009}
\pubyear{2009}
\jvolume{37}
\jissue{12}

@@

\batcave resembles an empirical Bayes approach, but is not really empirical Bayes.
In fact, variants are best described as sequential samples from a data generating distribution represented by the mutation profile.
Early mutations are classified with low error, while later mutations are classified with less certainty.
We show here that sequential updating of the dirichlet-multinomial posterior converges to the true data generating distribution using the set of early mutations.
Along with the assumption that tumor evolution is approximately neutral, this allows us to generate a tumor and site-specific that sharpens the classification and calibration of later mutations.
Thus, \batcave is best described as an algorithm trained on early mutations.

\bkmcomment{this needs more work}
We simulate tumors with three mutation profiles, and evaluate two real tumor data sets.
Among these performance is best with the most concentrated simulatin profile.
The intermediate and diffuse mutation profiles are designed to represent baseline or minimal profiles of Lung and Breast tumors respectively.
In fact, lung and breast tumors have heterogeneous mutation profiles with a wide range of levels of concentration.
The Acute Myeloid Leukemia we evaluated has a particularly diffuse profile, which is common for these tumors \cite{Alexandrov2019}.

@@

% \section{CONCLUSION}
% \bkmcomment{I looked at a bunch of NAR methods papers, and none include a conclusions section, so I moved it to the last paragraph of discussion.}

\section{Software Availability}
\bkmcomment{working on this today}

github for source code

zenodo for figures, simulations

zenodo for package at time of bioarxiv

@@

This work was supported by the National Science Foundation via Graduate Research Fellowship DGE-1143953 to BKM and \bkmcomment{RNG NIH}.
The authors would like to thank Professor Edward J. Bedrick for fruitful discussions about the statistical model.
